---
title: "Final Product"
author: "Palma"
date: "11/30/2021"
output: html_document
---

```{r, error=FALSE}
library(nflfastR)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(stringr)
library(lubridate)
library(tidymodels)        # for modeling
library(themis)            # for step functions for unbalanced data
library(doParallel)        # for parallel processing
library(stacks)            # for stacking models
library(naniar)            # for examining missing values (NAs)
library(moderndive)        # for King County housing data
library(vip)               # for variable importance plots
library(patchwork)         # for combining plots nicely
library(ranger)
library(readr)
library(ggimage)
library(xgboost)
library(kableExtra)
library(expss)
library(knitr)
library(skimr)
library(baguette) 
library(future)
```

```{r}
pbp_2018_2021 <- load_pbp(2018:2021)
nfl_qbr_weekly <- readr::read_csv("https://raw.githubusercontent.com/nflverse/espnscrapeR-data/master/data/qbr-nfl-weekly.csv")
nfl_qbr_weekly<-nfl_qbr_weekly %>% 
  filter(season==2018:2021)
```

# Intro to the "Two Minute Drill"

Show win probability .... what happens here that changes the outcome?

Show QB face plots

Show most successful plots

What will we be modeling / looking into?

# Intro Visualizations 

3 - 4 intro visualizations investigating interesting behavior....

specifically ones that directly have impications upon the outcome (score?)

```{r}
Two_min_drill <- pbp_2018_2021 %>% 
  filter(half_seconds_remaining<120, as.numeric(ms(drive_game_clock_start))<150) 
  
Two_min_drill %>% 
  group_by(game_id, drive) %>%
  select(game_id, drive_play_count, ydsnet, drive_game_clock_start, fixed_drive_result, name, posteam, week, drive_start_yard_line, wp) %>%
  mutate(td = ifelse(fixed_drive_result=='Touchdown', 1, 0),
  fg = ifelse(fixed_drive_result=='Field goal', 1, 0)) %>%
  mutate(score= ifelse(td+fg==1, 1,0))

```

```{r}
Two_min_drill

Two_min_drill %>% 
  group_by(game_id, drive) %>%
  mutate(rush_attempt = ifelse(is.na(rush_attempt), 0, rush_attempt)) %>% 
  mutate(pass_attempt = ifelse(is.na(pass_attempt), 0, pass_attempt)) %>%
  summarize(num_plays= n(), num_rush=sum(rush_attempt), num_pass= sum(pass_attempt))
  
```
```{r}
two_min_new <- Two_min_drill %>% 
  group_by(game_id, drive) %>%
  mutate(td = ifelse(fixed_drive_result=='Touchdown', 1, 0),
  fg = ifelse(fixed_drive_result=='Field goal', 1, 0), 
  score= ifelse(td+fg==1, 1,0)) %>%
  right_join(nfl_qbr_weekly, by = c("week" = "game_week", "posteam"="team_abb", "season" = "season")) %>% 
  ungroup()
```

```{r}
#two_min_new <- two_min_new[!is.na(two_min_new$passer), ] 
```

```{r}
two_min_new %>%
  select(passer, ydsnet, qbr_total, qbr_raw, game_id.x, headshot_href)
```

```{r}
two_min_by_drive <- 
  two_min_new  %>%
    group_by(drive, game_id.x) %>%
    mutate(run_plays = sum(rush_attempt, na.rm = TRUE), 
           pass_plays = sum(pass_attempt, na.rm = TRUE), 
           pass_tot_yds = sum(air_yards, na.rm = TRUE), 
           completion_perc = (1- sum(incomplete_pass, na.rm = TRUE) / pass_plays), 
           #penalties = sum(penalty, na.rm = TRUE), 
           tot_yds = sum(yards_gained, na.rm = TRUE),
           rush_yds_tot = sum(rushing_yards, na.rm = TRUE)) %>% 
           #rusher = as.character(rusher),
           #rusher = ifelse( is.na(rusher) == TRUE, getmode(rusher), rusher)) %>% 
     mutate(td = ifelse(fixed_drive_result=='Touchdown', 1, 0),
            fg = ifelse(fixed_drive_result=='Field goal', 1, 0), 
            score = ifelse(td+fg==1, 1,0)) %>%
    select(passer, qbr_raw, qbr_total, pass_tot_yds, tot_yds, ydsnet, rush_yds_tot, rusher, completion_perc, run_plays, pass_plays, drive_yards_penalized, tot_yds, drive_game_clock_start, td, fg, score, posteam, drive_start_yard_line, headshot_href, wp, season) %>% 
  mutate(yards_to_go_start= ifelse(str_extract(drive_start_yard_line, "[A-Z]+")== posteam, 100- parse_number(drive_start_yard_line), parse_number(drive_start_yard_line))) 
```

```{r}
drive_summary_data <- two_min_by_drive %>%
 arrange(game_id.x, drive) %>% 
 group_by(game_id.x) %>% 
  mutate(
    td = as.factor(td), 
    fg = as.factor(fg), 
    score = as.factor(score)
  ) %>%
 summarise_all(last)

drive_summary_data <- drive_summary_data %>% 
  mutate(passer= replace(passer, passer=="Aa.Rodgers", "A.Rodgers"))

drive_summary_data
```
```{r}
drive_summary_data %>% 
  summarize(num_td_drives= sum(td==1), num_fg_drives= sum(fg==1), num_drives=n(), td_success_perc= num_td_drives/num_drives, fg_success_perc= num_fg_drives/num_drives)
```

```{r}
drive_summary_data %>% 
  group_by(season) %>%
  summarize(season, num_td_drives= sum(td==1), num_fg_drives= sum(fg==1), num_drives=n(), td_success_perc= num_td_drives/num_drives, fg_success_perc= num_fg_drives/num_drives ) %>% 
  dplyr::slice(1) %>% 
  dplyr::rename("Number of TD's Scored"= num_td_drives) %>%
  dplyr::rename("Number of FG's Scored"= num_fg_drives) %>%
  dplyr::rename("Number of Drives"= num_drives) %>%
  dplyr::rename("TD Success Rate"= td_success_perc) %>%
  dplyr::rename("FG Success Rate"= fg_success_perc) %>%
  kbl() %>% 
  kable_paper("striped") %>% 
  add_header_above(c("2-minute Drills in the last 4 Seasons" = 6))  
```


```{r}
drive_summary_data %>%
  group_by(passer) %>% 
  summarize(num_successful= sum(score==1), num_2min_drives= n(), success_perc= num_successful/num_2min_drives, headshot=headshot_href) %>%
  arrange(desc(num_successful)) %>%  
  dplyr::slice(1) %>% 
  ggplot(aes(x=num_2min_drives, y= num_successful))+
  geom_text_repel(aes(label=passer)) +
  geom_image(aes(image= headshot), size=0.05, asp= 16/9)+
  labs(x= "Number of 2-minute drill drives",
       y= "Number of Successfull 2-minute drill drives",
       title= "Every QB 2-Minute Drill Efficiency from 2018-2021",
       caption = "Data: @nflfastR")+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 8)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 15))+
  theme_bw()
```

```{r}
drive_summary_data %>%
  ggplot(aes(x=qbr_total, y=wp, color=score))+ 
  geom_point()+
  labs(x="Total QBR for that week",
       y= "Current Winning Percentage during the Drive",
       title = "Does scoring(TD or FG) on a 2-Minute Drill Drive increase Win Percentage?",
       caption = "Data: @nflfastR",
       color= "Scored?")+
  scale_color_manual(labels = c("No", "Yes"), values = c("blue", "red"))
  
```





# Modeling 

- log regression

What did we model
Outcome
What can we do with this
How accurate is it
Make predictions on new people....or real cfb situations

Drawbacks / limitations

- decision tree

What did we model
Outcome
What can we do with this
How accurate is it
make predictions on new people... or real cfb situations

Drawbacks / limitations

# Conclusion 

<<<<<<< HEAD

=======
What did we uncover about two min drills?

When are they more successful? When are they less? What vars are the most important?
>>>>>>> df2a57f636984d74ab99d9deb37f4ec34c26f7a7
