---
title: "Work Doc - Ben"
author: "Wagner"
date: "11/16/2021"
output: html_document
---

For the projects, I want to give some suggested due dates:

Friday 11/19: 1. Create a GitHub project and add collaborators (I'll go through an example of this on Tuesday). 2. At least 1 data source obtained,  data read in to R, basic visualizations/summarizations to verify data is as expected. Start doing some data cleaning. 3. Outline of analysis plan with some of the steps completed.

Tuesday 11/23: 1. Read in and clean other two sources of data. 2. Finish a couple of the analysis steps and write them up completely. I recommend focusing on the "final product" first. 3. Create an RMarkdown file for the "behind-the-scenes" report and be sure to keep track of any detailed explanations of methods and code there. 

Thursday 12/2: 1. Complete most of the analyses and have a draft of the "final product". 2. Begin creating a presentation. 3. Create a detailed outline of what will go in the "behind-the-scenes". 

Tuesday 12/7: 1. "Final product" complete. 2. Finish the 15 minute presentation and practice it with your group. 

Thursday 12/9: 1. Submit "final product" on moodle. 2. Give presentations - we may need to use a bit of the lunch hour. 

Thursday 12/16: Submit "behind-the-scenes" (you can always submit early, if you'd like). 


```{r}
install.packages("nflfastR")
install.packages("nflreadr")
```

```{r}
library(nflfastR)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(lubridate)
pbp_2021 <- load_pbp(2021)
```

```{r}
Two_min_drill <- pbp_2021 %>% 
  filter(half_seconds_remaining<120, as.numeric(ms(drive_game_clock_start))<150) 
  
Two_min_drill %>% 
  group_by(game_id, drive) %>%
  summarise( drive_play_count, ydsnet, drive_game_clock_start, fixed_drive_result) %>%
  mutate(td = ifelse(fixed_drive_result=='Touchdown', 1, 0),
  fg = ifelse(fixed_drive_result=='Field goal', 1, 0))

```