---
title: "Work Doc - Ben"
author: "Wagner"
date: "11/16/2021"
output: html_document
---

For the projects, I want to give some suggested due dates:

Friday 11/19: 1. Create a GitHub project and add collaborators (I'll go through an example of this on Tuesday). 2. At least 1 data source obtained,  data read in to R, basic visualizations/summarizations to verify data is as expected. Start doing some data cleaning. 3. Outline of analysis plan with some of the steps completed.

Tuesday 11/23: 1. Read in and clean other two sources of data. 2. Finish a couple of the analysis steps and write them up completely. I recommend focusing on the "final product" first. 3. Create an RMarkdown file for the "behind-the-scenes" report and be sure to keep track of any detailed explanations of methods and code there. 

Thursday 12/2: 1. Complete most of the analyses and have a draft of the "final product". 2. Begin creating a presentation. 3. Create a detailed outline of what will go in the "behind-the-scenes". 

Tuesday 12/7: 1. "Final product" complete. 2. Finish the 15 minute presentation and practice it with your group. 

Thursday 12/9: 1. Submit "final product" on moodle. 2. Give presentations - we may need to use a bit of the lunch hour. 

Thursday 12/16: Submit "behind-the-scenes" (you can always submit early, if you'd like). 


```{r}
install.packages("nflfastR")
install.packages("nflreadr")
```

```{r}
library(nflfastR)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(lubridate)
library(stringr)
library(readr)
pbp_2021 <- load_pbp(2021)
nfl_qbr_weekly <- readr::read_csv("https://raw.githubusercontent.com/nflverse/espnscrapeR-data/master/data/qbr-nfl-weekly.csv")
nfl_qbr_weekly<-nfl_qbr_weekly %>% 
  filter(season==2021)
```

```{r}
Two_min_drill <- pbp_2021 %>% 
  filter(half_seconds_remaining<120, as.numeric(ms(drive_game_clock_start))<150) 
  
Two_min_drill %>% 
  group_by(game_id, drive) %>%
  select(game_id, drive_play_count, ydsnet, drive_game_clock_start, fixed_drive_result, name, posteam, week, drive_start_yard_line, wp) %>%
  filter(game_id == "2021_01_ARI_TEN", drive==11) %>% 
  mutate(td = ifelse(fixed_drive_result=='Touchdown', 1, 0),
  fg = ifelse(fixed_drive_result=='Field goal', 1, 0)) %>%
  mutate(yards_to_go_start= ifelse(str_extract(drive_start_yard_line, "[A-Z]+")== posteam, 100- parse_number(drive_start_yard_line), parse_number(drive_start_yard_line))) %>% 
  mutate(yards_to_go_start= parse_character(drive_start_line, na = c("", "NA"))) %>% 
  mutate(score= ifelse(td+fg==1, 1,0)) %>% 
  right_join(nfl_qbr_weekly, by = c("week" = "game_week", "posteam"="team_abb")) %>% 
  select(!game_id.y)
  # ggplot(aes(x= score, y=qbr_total))+
  # geom_boxplot()+
  # facet_wrap(vars(score))

```

```{r}
Two_min_drill

Two_min_drill %>% 
  group_by(game_id, drive) %>%
  mutate(rush_attempt = ifelse(is.na(rush_attempt), 0, rush_attempt)) %>% 
  mutate(pass_attempt = ifelse(is.na(pass_attempt), 0, pass_attempt)) %>%
  summarize(num_plays= n(), num_rush=sum(rush_attempt), num_pass= sum(pass_attempt))
  
```

